LAGOS lake depth

### 2020-11-23 figures/01_geometry.Rmd
prototype slope alternative scatter plots

```{r 01_geometry_base, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
theme_opts <- cowplot::theme_cowplot() + 
  theme(legend.position = "none", 
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 11))
text_size <- 3

@@

gg_dist_base <- st_drop_geometry(dt) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::filter(dist_viscenter >= dist_deepest) %>%
  dplyr::filter(dist_deepest < 3200) %>%
  ggplot(aes(x = dist_deepest, y = dist_viscenter)) + 
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = 2550, y = 300, label = "R^2 == 0.80",
                                parse = TRUE, size = text_size) +
  annotate("text", x = 2100, y = 150, 
           label = get_formula(
             lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))), 
           size = text_size) +
  labs(color = "") +
  xlab("Deepest point distance (m)") +
  ylab("Distance to \n lake center (m)") +
  xlim(0, 3000) +
  theme_opts

@@

gg_slope_base <- ggplot(data = test) + 
  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt), size = 1) +
  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt), method = "lm", se = FALSE) +
  geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
  alpha = 1, fill = "white") +
  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
                                parse = TRUE, size = text_size) +
  annotate("text", x = 0.34, y = 0.01, 
           label = get_formula(
             lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
           size = text_size) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "") +
  theme_opts

@@

>>>>

```{r 01_geometry_base, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
# setwd("figures")
theme_opts <- cowplot::theme_cowplot() + 
  theme(legend.position = "none", 
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 11))
text_size <- 3

@@

rsq <- round(broom::glance(lm(dist_viscenter ~ dist_deepest, data = dt))$r.squared, 2)
gg_dist_base <- st_drop_geometry(dt) %>%
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::filter(dist_viscenter >= dist_deepest) %>%
  dplyr::filter(dist_deepest < 3200) %>%
  ggplot(aes(x = dist_deepest, y = dist_viscenter)) + 
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = 2550, y = 400, label = paste0("R^2 == ", rsq),
                                parse = TRUE, size = text_size-0.7) +
  annotate("label", x = 1800, y = 150, 
           label = get_formula(
             lm(dist_viscenter ~ dist_deepest, data = st_drop_geometry(dt))), 
           size = text_size-0.85, fill = "white", label.size = NA) +
  labs(color = "") +
  xlab("Deepest point distance (m)") +
  ylab("Distance to \n lake center (m)") +
  xlim(0, 3000) +
  theme_opts

@@

rsq <- round(broom::glance(lm(nearshore_slope_mean ~ inlake_slope_pnt, data = test))$r.squared, 2)
gg_slope_base <- ggplot(data = test) + 
  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt), size = 1) +
  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt), method = "lm", se = FALSE) +
  geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
  alpha = 1, fill = "white") +
  annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
                                parse = TRUE, size = text_size-0.7) +
  annotate("label", x = 0.27, y = 0.01, 
           label = get_formula(
             lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
           size = text_size-0.85, fill = "white", label.size = NA) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "") +
  theme_opts

@@

```{r 01_dist_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
# fig 2A (dist) alternatives
#   gg_dist_base <- dist_viscenter ~ dist_deepest
#   gg_dists_base <- dist_viscenter ~ dists_deepest
dt$dists_deepest <- sapply(dt$dists_deepest, as.numeric)
rsq <- round(broom::glance(lm(dist_viscenter ~ dists_deepest, data = dt))$r.squared, 2)

gg_dists_base <- dt %>%
  # dplyr::filter(!is.na(shape_class)) %>%
  dplyr::filter(dist_viscenter >= dist_deepest) %>%
  dplyr::filter(dist_deepest < 3200) %>%
  ggplot(aes(x = dists_deepest, y = dist_viscenter)) + 
  geom_point(size = 1) +
  geom_smooth(method = "lm", se = FALSE) +
  annotate("text", x = 1800, y = 400, label = paste0("R^2 == ", rsq),
                                parse = TRUE, size = text_size-0.7) +
  annotate("label", x = 1200, y = 150, 
           label = get_formula(
             lm(dist_viscenter ~ dists_deepest, data = dt)), 
           size = text_size-0.85, fill = "white", label.size = NA) +
  labs(color = "") +
  xlab("Deepest point distance (m)") +
  ylab("Distance to \n lake center (m)") +
  xlim(0, 2000) +
  theme_opts

cowplot::plot_grid(gg_dist_base, gg_dists_base)
```

```{r 01_slope_compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4.8, fig.height=2.3}
# fig 2B (slope) alternatives
#   gg_slope_base <- nearshore_slope_mean ~ inlake_slope_pnt
#   gg_slope_* <- nearshore_slope* ~ inlake_slope*

inlake_slope_alternatives     <- c("inlake_slope_pnt", "inlake_slope_pnts",
                                   "inlake_slope_mean",
                                   "inlake_slope_online_mean",
                                   "inlake_slopes_online_mean")
nearshore_slope_alternatives  <- c("nearshore_slope_mean",
                                   "nearshore_slope_online_mean",
                                   "nearshore_slopes_online_mean")
slope_alternatives   <- setNames(data.frame(expand.grid(
  inlake_slope_alternatives,
  nearshore_slope_alternatives,
  stringsAsFactors = FALSE
  )), c("inlake_slope", "nearshore_slope"))

gg_slope_alternative <- function(nearshore_var, inlake_var, dt){
  # dt <- test
  # nearshore_var <- "nearshore_slope_mean"
  # inlake_var <- "inlake_slope_pnt"
  rsq <- round(broom::glance(
    lm(as.formula(paste0(nearshore_var, " ~ ", inlake_var)), data = dt)
    )$r.squared, 2)
  res <-  ggplot(data = dt) + 
    geom_point(aes(.data[[nearshore_var]], .data[[inlake_var]]), size = 1) +
    geom_smooth(aes(.data[[nearshore_var]], .data[[inlake_var]]), 
                method = "lm", se = FALSE) +
    geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
    alpha = 1, fill = "white") +
    annotate("text", x = 0.34, y = 0.05, label = paste0("R^2 == ", rsq),
                                  parse = TRUE, size = text_size-0.7) +
    # annotate("label", x = 0.27, y = 0.01, 
    #          label = get_formula(
    #            lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
    #          size = text_size-0.85, fill = "white", label.size = NA) +
    # geom_abline(slope = 1, intercept = 0) +
    xlab(nearshore_var) +
    ylab(inlake_var) +
    labs(color = "") +
    theme_opts

  res
}

gg_slope_list <- lapply(seq_len(nrow(slope_alternatives)), function(i)
  gg_slope_alternative(slope_alternatives$nearshore_slope[i],
                     slope_alternatives$inlake_slope[i], test)
)

gg_slope_list[c(1:4, 6:9, 11:14)] <- lapply(
  gg_slope_list[c(1:4, 6:9, 11:14)], function(x) x +
     theme(axis.title.x = element_blank(), 
           axis.text.x = element_blank()))

gg_slope_list[c(6:15)] <- lapply(
  gg_slope_list[c(6:15)], function(x) x +
     theme(axis.title.y = element_blank(), 
           axis.text.y = element_blank()))

cowplot::plot_grid(plotlist = gg_slope_list, 
                   nrow = 5, byrow = FALSE, 
                   rel_heights = c(rep(0.6, 4), 1,
                                   rep(0.6, 4), 1,
                                   rep(0.6, 4), 1))
```
